### 2023-05-31 manuscript/chapters/1_intro.qmd
yay for more writing !

::: hidden
{{< include ../macros.tex >}}
:::

@@

This model can intuitively be seen as the long time limit of a network of spiking neurons, in which neurons emit current spikes (modeled as $\delta$-functions) at some time-variable rate $\phi(h_i(t))$. The convergence is formally proven in @schmutz2023convergence.

@@

In the case that $\vec F_i = (F_{1,i},\cdots,F_{p,i}) \in \Rp$ sample from some $p$-dimensional distribution $\rho$, the vectors $\vec F_i$ define a "natural embedding" of neuron $i$ in $\Rp$. When the number of neurons increases, the numeric density of neurons in the embedding approaches the actual probability distribution $\rho$, and we can intuitively see the convergence towards a connectivity kernel $w(\vec z, \vec y) = \sum_{\mu=1}^p z_\mu y_\mu$.

@@

$$
\partial_t \tilde h(t, \alpha) = -\tilde h(t, \alpha) + \int_{\mathbb{R}^p} \tilde w(\alpha, \beta) \phi(\tilde h(t, \beta)) \tilde \rho(\mathrm d \beta)
$$ {#eq-nf1d}

@@

With this in mind, we can try to understand why simulating a neural field in $[0,1]$ might be difficult. Suppose we take bins in the $\Rp$ space, and map each bin to a random position in $[0,1]$. Intuively, this won't work (in the sense that the dynamics in $[0,1]$ will not be the same as the dynamics in $\Rp$), because the hypothesis of continuity is not longer true : the mapping is random, so the nearby populations in $[0,1]$ will no longer have similar potentials, and the error of the mean-field approximation will no longer vanish as we take increasingly many bins.

@@

In some sense, we would like to find a mapping that converses "locality" : populations that are close in $\Rp$ should also be close in $[0,1]$, or at least there shouldn't be "too many large discontinuities". @sec-mapping is dedicated to the discussion of these mappings, and gives various numerical demonstrations of the resulting dynamics.

>>>>

::: {.content-hidden}
{{< include ../macros.tex >}}
:::

@@

This model can intuitively be seen as the $N \rightarrow \infty$ limit of a network of spiking neurons, in which neurons emit current spikes (modeled as $\delta$-functions) at some time-variable rate $\phi(h_i(t))$. The convergence is formally proven in @schmutz2023convergence.

@@

In the case that $\vec F_i = (F_{1,i},\cdots,F_{p,i}) \in \Rp$ sample from some $p$-dimensional distribution $\rho$, the vectors $\vec F_i$ define a "natural embedding" of neuron $i$ in $\Rp$. When the number of neurons increases, the numeric density of neurons in the embedding approaches the actual probability distribution $\rho$, and we can intuitively see the convergence towards a connectivity kernel $w(\vec z, \vec y) = \sum_{\mu=1}^p z_\mu g_\mu(y_\mu)$.

@@

$$
\partial_t \tilde h(t, \alpha) = -\tilde h(t, \alpha) + \int_{[0,1]} \tilde w(\alpha, \beta) \phi(\tilde h(t, \beta)) \tilde \rho(\mathrm d \beta)
$$ {#eq-nf1d}

@@

With this in mind, we can try to understand why simulating a neural field in $[0,1]$ might be difficult. Suppose we take bins in the $\Rp$ space, and map each bin to a random position in $[0,1]$. Intuitively, this won't work (in the sense that the dynamics in $[0,1]$ will not be the same as the dynamics in $\Rp$), because the hypothesis of continuity is not longer true : the mapping is random, so the nearby populations in $[0,1]$ will no longer have similar potentials, and the error of the mean-field approximation will no longer vanish as we take increasingly many bins.

@@

In some sense, we would like to find a mapping that conserves "locality" : populations that are close in $\Rp$ should also be close in $[0,1]$, or at least there shouldn't be "too many large discontinuities". @sec-mapping is dedicated to the discussion of these mappings, and gives various numerical demonstrations of the resulting dynamics.

### 2023-05-31 manuscript/chapters/2_neural_fields.qmd
yay for more writing !

# A simple model for a neural field in $\mathbb R^2$

## Building up a neural field in $\mathbb R^2$

p=1 then p=2, show simulations

@@

@@

overlaps as an approximation of the "dynamics" (latent space), in the case phi=linear

@@

@@

>>>>

# A simple model for a neural field in 2D

## Networks of neurons to neural fields

no external current,

general low-rank neural fields, F and G notation

$$
J_{ij} = \sum_{\mu=1}^p F_{\mu,i} G_{\mu,j}
$$

show how the embedding becomes smooth

study of fixed points

- $h_i = \xi_{\mu,i} \quad \forall \mu = 1,\cdots,p$ (stable)
- $h_i=0$ (unstable)

the fixed points also become more and more stable as N->oo (study of stability with eigenvalues)

## Building up to a neural field in $\mathbb R^2$

Introduce the neural field in 1D, then 2D

Neurons are at position $\vec z = (z_1, \cdots, z_p) \in \mathbb{R}^p$, distributed according to the distribution $\frac{\mathrm{exp}(\frac12\sum_{\mu=1}^p z_\mu^2)}{(2\pi)^{p/2}} \mathrm{d} z_1 \cdots \mathrm d z_p = \rho(\mathrm d \vec z)$

The RNN potential becomes $h(t, \vec z)$ and evolves according to

$$
\begin{aligned}
\partial_t h(t, \vec z) &= -h(t, \vec z) + \int_{\mathbb{R}^p} w(\vec z, \vec y) \phi(h(t, \vec y)) \rho(\mathrm d \vec y),
\quad
w(\vec z, \vec y) = \sum_{\mu=1}^p \tilde \phi (y_\mu) z_\mu \\
&= -h(t, \vec z) + \sum_{\mu=1}^p z_\mu m_\mu(t)
\end{aligned}
$$

study of fixed points (and convergence of the fixed points)

simulation strategies : grid and sampling

@@

### Overlap variables

networks of neurons :

$$
m_\mu(t) = \frac 1N \sum_{i=1}^N \tilde \phi(\xi_{\mu,i}) \phi(h_i(t))
$$

neural field :

$$
m_\mu(t) = \int_{\mathbb R^p} \tilde \phi(y_\mu) \phi(h(t,\vec y)) \rho(\mathrm d \vec y)
$$

### Low-rank neural fields as $p$-dimensional closed systems

@@

overlaps as an approximation of the “dynamics” (latent space), in the case phi=linear

The dynamics of $h(t, \vec z)$ are in a subsystem of dimension $p$ spanned by the ONB of functions $\{e_\mu(\vec z) = z_\mu | \mu=1,\cdots,p\}$, with the scalar product $\langle f, g \rangle = \int_{\mathbb R^p} f(\vec y) g(\vec y) \rho(\mathrm d \vec y)$.

We decompose $h(t, \vec z) = h^\perp(t, \vec z) + \sum_{\mu=1}^p \kappa_\mu(t) z_\mu$, and write the system of equations for $\mu=1,\cdots,p$.

$$
\dot \kappa_\mu(t) = -\kappa_\mu(t) + \int_{\mathbb{R}^p} \tilde\phi(y_\mu) \phi(h(t, \vec y)) \rho(\mathrm d \vec y) = -\kappa_\mu(t) + m_\mu(t)
$$

with initial conditions

$$
\kappa_\mu(0) = \int_{\mathbb{R}^p} y_\mu h(0, \vec y) \rho(\mathrm d \vec y)
$$

and the orthogonal component evolves according to

$$
h^\perp(t, \vec z) = h^\perp(0, \vec z) \mathrm e^{-t},
\quad
h^\perp(0, \vec z) = h(0, \vec z) - \sum_{\mu=1}^p \kappa_\mu(0) z_\mu
$$

### Dynamics in the overlap latent space

@@

Relation between $p$-dimensional closed system and neural field equation in $\mathbb R^p$

$$
\begin{aligned}
\partial_t h(t, \vec z) &= \partial_t h^\perp(t, \vec z) + \sum_{\mu=1}^p z_\mu \dot \kappa_\mu(t) \\
&= -h^\perp(t, \vec z) - \sum_{\mu=1}^p z_\mu \kappa_\mu(t) + \sum_{\mu=1}^p z_\mu m_\mu(t) \\
&= -h(t, \vec z) + \sum_{\mu=1}^p z_\mu m_\mu(t)
\end{aligned}
$$

we use the latent space to assert dynamics are the same

@@

recurrent current picture
